Automatic prostate cancer detection through \acs*{dce}-\acs*{mri} images: all you need is a good normalization

### 2016-09-01 01_introduction/introduction.tex
Add full title page

The paper is organized as follows:
Section~\ref{sec:norm} details our normalization strategy for \ac{dce}-\ac{mri} data.
Quantitative and semi-quantitative methods are summarized in Sect.\,\ref{sec:stateart} with insights about their implementations.
Section~\ref{sec:materials} gives information about the dataset used and provided source code.
Experiments and results to answer the previous stated challenges are reported in Sect.\,\ref{sec:experiments} while discussed in Sect.\,\ref{sec:discussions}, followed by a concluding section.
%Section~\ref{sec:methods} outlines our normalization strategy (Section~\ref{sec:norm}) as well as specificity regarding the state-of-the-art methods used for comparison (Section~\ref{sec:stateart}).
%The dataset, experiments, and results are reported in Section~\ref{sec:experiments} while discussed in Section~\ref{sec:discussions} followed by a concluding section.

>>>>

The paper is organized as follows:
Section~\ref{sec:data} gives information about the dataset used and provided source code.
Section~\ref{sec:norm} details our normalization strategy for \ac{dce}-\ac{mri} data.
Quantitative and semi-quantitative methods are summarized in Sect.\,\ref{sec:stateart} with insights about their implementations.
Experiments and results to answer the previous stated challenges are reported in Sect.\,\ref{sec:experiments} while discussed in Sect.\,\ref{sec:discussions}, followed by a concluding section.
%Section~\ref{sec:methods} outlines our normalization strategy (Section~\ref{sec:norm}) as well as specificity regarding the state-of-the-art methods used for comparison (Section~\ref{sec:stateart}).
%The dataset, experiments, and results are reported in Section~\ref{sec:experiments} while discussed in Section~\ref{sec:discussions} followed by a concluding section.

### 2016-09-01 02_methods/methods.tex
Add full title page

\section{Methods}\label{sec:method}

>>>>

\section{Material and Methods}\label{sec:method}

\subsection{\ac{mri} data}\label{sec:data}

The multi-parametric \ac{mri} data are acquired from a cohort of patients with higher-than-normal level of \ac{psa}.
The acquisition is performed using a 3T whole body \ac{mri} scanner (Siemens Magnetom Trio TIM, Erlangen, Germany) using sequences to obtain \ac{t2w}-\ac{mri}, \ac{dce}-\ac{mri} and \ac{dw}-\ac{mri}.
Aside of the \ac{mri} examination, these patients also have undergone a guided-biopsy.
The dataset is composed of a total of 20 patients of which 18 patients have biopsy proven \ac{cap} and 2 patients are ``healthy'' with negative biopsies.
Therefore, 13 patients have a \ac{cap} in the \ac{pz}, 3 patients have \ac{cap} in the \ac{cg}, 2 patients have invasive \ac{cap} in both \ac{pz} and \ac{cg} and finally 2 patients are considered as ``healthy''.
An experienced radiologist has segmented the prostate organ --- on \ac{t2w}-\ac{mri} and \ac{dce}-\ac{mri} --- as well as the prostate zones (i.e., \ac{pz} and \ac{cg}) and \ac{cap} on the \ac{t2w}-\ac{mri}.

A \SI{3}{\mm} slice fat-suppressed \ac{t2w} fast spin-echo sequence (\ac{tr}/\ac{te}/\ac{etl}: \SI{3400}{\ms}/\SI{85}{\ms}/13) is used to acquire images in sagittal and oblique coronal planes, the latter planes being orientated perpendicular or parallel to the prostate \ac{pz} – rectal wall axis.
Three-dimensional \ac{t2w} fast spin-echo (\ac{tr}/\ac{te}/\ac{etl}: \SI{3600}{\ms}/\SI{143}{\ms}/109, slice thickness: \SI{1.25}{\mm}) images are then acquired in an oblique axial plane.
The nominal matrix and \ac{fov} of the 3D \ac{t2w} fast spin-echo images are $320 \times 256$ and $280 \times 240$ mm\textsuperscript{2}, respectively, thereby affording sub-millimetric pixel resolution within the imaging plane.

\ac{dce}-\ac{mri} is performed using a fat suppressed 3D T$_1$ VIBE sequence (\ac{tr}/\ac{te}/Flip angle: \SI{3.25}{\ms}/\SI{1.12}{\ms}/\SI{10}{\degree}; Matrix: $256 \times 192$; \ac{fov}: $280 \times 210$ (with 75\% rectangular \ac{fov}); slab of 16 partitions of \SI{3.5}{\mm} thickness; temporal resolution: \SI{6}{\s}/slab over approximately \SI{5}{\minute}).
A power injector (Medrad, Indianola, USA) is used to provide a bolus injection of Gd-DTPA (Dotarem, Guerbet, Roissy, France) at a dose of \SI{0.2}{\ml} Gd-DTPA/kg of body weight.

These \ac{dce}-\ac{mri} sequences are resampled using the spatial information of the \ac{t2w}-\ac{mri} and missing data are interpolated using a linear interpolation.
The volumes of the \ac{dce}-\ac{mri} dynamic are rigidly registered, to remove any patient motion during the acquisition.
Furthermore, a non-rigid registration is performed between the \ac{t2w}-\ac{mri} and \ac{dce}-\ac{mri} in order to propagate the prostate zones and \ac{cap} ground-truths.

The resampling is implemented in C++ using the Insight Segmentation and Registration Toolkit~\citep{ibanez2005itk}.
The implementation of the registration (C++), normalization (Python), and classification pipeline (Python) are publicly available on GitHub\footnote{\url{https://github.com/I2Cvb/lemaitre-2016-nov/tree/master}}~\citep{lemaitre2016github}.
The data used for this work are also publicly available\footnote{\url{https://zenodo.org/record/61163}}~\citep{lemaitre2016dce}.

### 2016-09-01 03_experiments/experiments.tex
Add full title page

\section{Materials}\label{sec:materials}

\subsection{Data}\label{sec:data}

The multi-parametric \ac{mri} data are acquired from a cohort of patients with higher-than-normal level of \ac{psa}.
The acquisition is performed using a 3T whole body \ac{mri} scanner (Siemens Magnetom Trio TIM, Erlangen, Germany) using sequences to obtain \ac{t2w}-\ac{mri}, \ac{dce}-\ac{mri} and \ac{dw}-\ac{mri}.
Aside of the \ac{mri} examination, these patients also have undergone a guided-biopsy.
The dataset is composed of a total of 20 patients of which 18 patients have biopsy proven \ac{cap} and 2 patients are ``healthy'' with negative biopsies.
Therefore, 13 patients have a \ac{cap} in the \ac{pz}, 3 patients have \ac{cap} in the \ac{cg}, 2 patients have invasive \ac{cap} in both \ac{pz} and \ac{cg} and finally 2 patients are considered as ``healthy''.
An experienced radiologist has segmented the prostate organ --- on \ac{t2w}-\ac{mri} and \ac{dce}-\ac{mri} --- as well as the prostate zones (i.e., \ac{pz} and \ac{cg}) and \ac{cap} on the \ac{t2w}-\ac{mri}.

A \SI{3}{\mm} slice fat-suppressed \ac{t2w} fast spin-echo sequence (\ac{tr}/\ac{te}/\ac{etl}: \SI{3400}{\ms}/\SI{85}{\ms}/13) is used to acquire images in sagittal and oblique coronal planes, the latter planes being orientated perpendicular or parallel to the prostate \ac{pz} – rectal wall axis.
Three-dimensional \ac{t2w} fast spin-echo (\ac{tr}/\ac{te}/\ac{etl}: \SI{3600}{\ms}/\SI{143}{\ms}/109, slice thickness: \SI{1.25}{\mm}) images are then acquired in an oblique axial plane.
The nominal matrix and \ac{fov} of the 3D \ac{t2w} fast spin-echo images are $320 \times 256$ and $280 \times 240$ mm\textsuperscript{2}, respectively, thereby affording sub-millimetric pixel resolution within the imaging plane.

\ac{dce}-\ac{mri} is performed using a fat suppressed 3D T$_1$ VIBE sequence (\ac{tr}/\ac{te}/Flip angle: \SI{3.25}{\ms}/\SI{1.12}{\ms}/\SI{10}{\degree}; Matrix: $256 \times 192$; \ac{fov}: $280 \times 210$ (with 75\% rectangular \ac{fov}); slab of 16 partitions of \SI{3.5}{\mm} thickness; temporal resolution: \SI{6}{\s}/slab over approximately \SI{5}{\minute}).
A power injector (Medrad, Indianola, USA) is used to provide a bolus injection of Gd-DTPA (Dotarem, Guerbet, Roissy, France) at a dose of \SI{0.2}{\ml} Gd-DTPA/kg of body weight.

These \ac{dce}-\ac{mri} sequences are resampled using the spatial information of the \ac{t2w}-\ac{mri} and missing data are interpolated using a linear interpolation.
The volumes of the \ac{dce}-\ac{mri} dynamic are rigidly registered, to remove any patient motion during the acquisition.
Furthermore, a non-rigid registration is performed between the \ac{t2w}-\ac{mri} and \ac{dce}-\ac{mri} in order to propagate the prostate zones and \ac{cap} ground-truths.
The resampling is implemented in C++ using the Insight Segmentation and Registration Toolkit~\citep{ibanez2005itk}.

\subsection{Implementation}

The implementation of the registration (C++), normalization (Python), and classification pipeline (Python) are publicly available on GitHub\footnote{\url{https://github.com/I2Cvb/lemaitre-2016-nov/tree/master}}~\citep{lemaitre2016github}.
The data used for this work are also publicly available\footnote{\url{https://zenodo.org/record/61163}}~\citep{lemaitre2016dce}.

\section{Experiments and results}\label{sec:experiments}

>>>>

%\section{Materials}\label{sec:materials}

\section{Results}\label{sec:experiments}

### 2016-09-01 full_title.tex
Add full title page

\documentclass[review]{elsarticle}

@@

\journal{Medical Image Analysis}

@@

%% Numbered
%\bibliographystyle{00_files_system/model1-num-names}

@@

%% Harvard
\bibliographystyle{00_files_system/model2-names}\biboptions{authoryear}

@@

% Add authors and the different addresses
\author[label1,label3]{Guillaume~Lema\^itre\corref{cor1}}
\ead{g.lemaitre58@gmail.com}
\author[label3]{Robert~Mart\'i}
\author[label1]{Mojdeh~Rastgoo}
\author[label1]{Joan~Massich}
%\ead{marly@eia.udg.edu}
\author[label3]{Jordi~Freixenet}
%\ead{jordif@eia.udg.edu}
\author[label5]{Anke Meyer-Baese}
%\ead{ameyerbaese@fsu.edu}
\author[label4]{Joan~C.~Vilanova}
%\ead{kvilanova@comg.cat}
%\author[label2]{Paul~M.~Walker}
%\ead{pwalker@u-bourgogne.fr}
\author[label1,label6]{Fabrice~Meriaudeau}
%\ead{fabrice.meriaudeau@u-bourgogne.fr}
\address[label1]{\scriptsize LE2I UMR6306, CNRS, Arts et M\'etiers, Univ. Bourgogne Franche-Comt\'e, 12 rue de la Fonderie, 71200 Le Creusot, France}
%\address[label2]{\scriptsize LE2I UMR6306, CNRS, Arts et M\'etiers, Univ. Bourgogne Franche-Comt\'e, Avenue Alain Savary, 21000 Dijon, France}
\address[label3]{\scriptsize ViCOROB, Universitat de Girona, Campus Montilivi, Edifici P4, 17071 Girona, Spain}
\address[label4]{\scriptsize Department of Magnetic Resonance, Cl\'inica Girona, Lorenzana 36, 17002 Girona, Spain}
\address[label5]{\scriptsize Department of Scientific Computing, 400 Dirac Science Library, Florida State University, Tallahassee, FL 32306, United States}
\address[label6]{\scriptsize CISIR, Electrical \& Electronic Engineering Department, Universiti Teknologi Petronas, 32610 Seri Iskandar, Perak, Malaysia}
\cortext[cor1]{Corresponding author.}

% Start the abstract
\begin{abstract}
\ac{mpmri} plays a major role in the detection of \ac{cap} and is commonly used to the design of new \ac{cad} systems.
\ac{dce}-\ac{mri} is one of the modality regularly used in \ac{mpmri} \ac{cad} system.
Pharmocokinetric parameters are extracted from the \ac{dce}-\ac{mri} sequences and later used to discriminate cancerous from healthy tissue.
However, some inter-patients variations occur during the data acquisition leading to estimation errors of the pharmacokinetic parameters.
Therefore, we propose a fully automatic normalization method for \ac{dce}-\ac{mri} that reduces the inter-patient variability of the data.
The benefit and simplicity of our approach is shown by detecting cancerous voxel from healthy voxel using the normalized enhanced \ac{dce}-\ac{mri} signal and comparing with the state-of-the-art quantitative and semi-quantitative methods.
Additionally, we show that using this normalization approach in conjunction with the quantitative methods improves the classification performance of most of the models.
The best classification performance to detect \ac{cap} is obtained using the whole \ac{dce}-\ac{mri} normalized signal, reaching an \ac{auc} of $0.666\ (\pm 0.154)$ and outperforming the quantitative approaches.
\end{abstract}

% Give the different keywords
\begin{keyword}
\acs*{dce}-\acs*{mri}\sep prostate cancer\sep normalization\sep classification\sep quantification
\end{keyword}

@@

% Reset all the acronyms
\acresetall
\linenumbers
% Include now the different files
\input{01_introduction/introduction}
\input{02_methods/methods}
\input{03_experiments/experiments}
\input{04_discussions/discussions}
\input{05_conclusions/conclusions}
\input{06_appendix/appendix}

% Include the bibliography
\section*{References}
\bibliography{00_files_system/literature}

\input{07_tables/tables}
\input{08_figures/figures}

@@

>>>>

\documentclass[review, 12pt]{elsarticle}

@@

\journal{Journal of Magnetic Resonance Imaging}

@@

%% Numbered
\bibliographystyle{00_files_system/model1-num-names}

@@

%% Harvard
%\bibliographystyle{00_files_system/model2-names}\biboptions{authoryear}

@@

% Add authors and the different addresses
\author[label1,label3]{Guillaume~Lema\^itre\corref{cor1}, MS}
\ead{g.lemaitre58@gmail.com}
\author[label3]{Robert~Mart\'i, PhD}
\author[label1]{Mojdeh~Rastgoo, PhD}
\author[label1]{Joan~Massich, PhD}
%\ead{marly@eia.udg.edu}
\author[label3]{Jordi~Freixenet, PhD}
%\ead{jordif@eia.udg.edu}
\author[label5]{Anke Meyer-Baese, PhD}
%\ead{ameyerbaese@fsu.edu}
\author[label4]{Joan~C.~Vilanova, MD}
%\ead{kvilanova@comg.cat}
%\author[label2]{Paul~M.~Walker}
%\ead{pwalker@u-bourgogne.fr}
\author[label1,label6]{Fabrice~Meriaudeau, PhD}
%\ead{fabrice.meriaudeau@u-bourgogne.fr}
\address[label1]{\scriptsize LE2I UMR6306, CNRS, Arts et M\'etiers, Univ. Bourgogne Franche-Comt\'e, 12 rue de la Fonderie, 71200 Le Creusot, France}
%\address[label2]{\scriptsize LE2I UMR6306, CNRS, Arts et M\'etiers, Univ. Bourgogne Franche-Comt\'e, Avenue Alain Savary, 21000 Dijon, France}
\address[label3]{\scriptsize ViCOROB, Universitat de Girona, Campus Montilivi, Edifici P4, 17071 Girona, Spain}
\address[label4]{\scriptsize Department of Magnetic Resonance, Cl\'inica Girona, Lorenzana 36, 17002 Girona, Spain}
\address[label5]{\scriptsize Department of Scientific Computing, 400 Dirac Science Library, Florida State University, Tallahassee, FL 32306, United States}
\address[label6]{\scriptsize CISIR, Electrical \& Electronic Engineering Department, Universiti Teknologi Petronas, 32610 Seri Iskandar, Perak, Malaysia}
\cortext[cor1]{Corresponding author: Guillaume Lema\^itre, 12 rue de la Fonderie, 71200 Le Creusot, +33761104782}

\begin{abstract}
\begin{description}
  \item[Acknowledgements] None.
  \item[Grant support] Guillaume Lema\^itre is supported by the Generalitat de Catalunya (grant nb. FI-DGR2012) and partly by the Mediterranean Office for Youth (grant nb. 2011/018/06).
  \item[Running title] All you need is a good normalization
\end{description}
\end{abstract}

@@

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:
